Generative AI for image reconstruction in Intensity Interferometry: a first attempt

### 2025-04-30 Phase_II/draft/conclusion.tex
edits

However, II faces a significant drawback: it captures only the magnitude of the signal via photon correlation, resulting in a loss of phase information. This challenge of phase retrieval in II has been effectively addressed through machine learning techniques, particularly with conditional Generative Adversarial Networks (cGAN). Our study demonstrates that applying cGAN to II data successfully recovers the size, shape, and brightness distribution of a fast-rotating star. Evaluations based on image moments—specifically, the monopole, second, and third-order moments—support the effectiveness of cGAN in achieving accurate image reconstruction from a one-night simulation of II using six baselines. A critical factor in the reconstruction process is the extent of Fourier plane coverage, which depends on the number of available telescopes and the total observing time. With a full night of observation using four telescopes, the brightness distribution could be reconstructed with even greater precision. Future work might explore different observatory layouts to assess their impact on image reconstruction quality, such as integrating the Southern Cherenkov Telescope Array (CTA) with MAGIC.

While the results of this study highlight the significant potential of machine learning—particularly cGAN—for image reconstruction in II, several aspects require further refinement. First, detector efficiencies, which impact the signal-to-noise ratio (SNR) of actual observational data, have not yet been incorporated; addressing these factors will be crucial for more accurate SNR estimation. Second, exploring and comparing alternative methods for image generation could reveal approaches that outperform cGAN in reconstructing stellar images with II. Third, experimenting with different loss functions could provide additional insights into reconstruction quality. Although further testing is needed to refine the GAN and enhance its robustness and reliability, our findings suggest that machine learning is a promising approach for phase reconstruction in II.

>>>>

Intensity Interferometry (II) is re-emerging as a promising technique to overcome the challenges associated with Michelson Interferometry in the optical wavelength range.  However, compared to radio-interferometry II faces a significant drawback: photon correlation captures only the magnitude of the interferometric signal, resulting in a loss of phase information.

This work addresses the challenge of phase retrieval in II using a machine-learning technique, specifically a conditional Generative Adversarial Networks (cGAN). Our study demonstrates that applying cGAN to II data successfully recovers the size, shape, and brightness distribution of a fast-rotating star. Evaluations based on image moments—specifically, the monopole, second, and third-order moments—support the effectiveness of cGAN in achieving accurate image reconstruction from a simulation of II from a single site with four telescopes.

While the results of this study highlight the significant potential of machine learning, and in particular cGAN, for image reconstruction in II, several aspects require further refinement. First, an important factor in the reconstruction process is the extent of Fourier plane coverage, which depends on the number of available telescopes and the total observing time. With more than four telescopes, more complicated systems could be reconstructed. Future work might explore different observatory layouts to assess their impact on image reconstruction quality.  Second, detector efficiencies, which impact the signal-to-noise ratio (SNR) of actual observational data, have not yet been incorporated; addressing these factors will be crucial for more accurate SNR estimation.  Third, exploring and comparing alternative methods for image generation could reveal approaches that outperform cGAN in reconstructing stellar images with II. Fourth, experimenting with different loss functions could provide additional insights into reconstruction quality. Although further testing is needed to refine the GAN and enhance its robustness and reliability, our findings suggest that machine learning is a promising approach for phase reconstruction in II.

### 2025-04-30 Phase_II/draft/main.tex
edits

\title{Image reconstruction of stellar objects with Intensity Interferometry using Generative Adversarial Networks}
\date{}

@@

\begin{abstract}
In modern astronomy, simulations of Intensity Interferometry are making significant strides in achieving high-precision resolution of stellar objects at optical wavelengths. Despite these advancements, phase retrieval remains a major challenge due to the nature of photon correlation. This paper explores the application of a conditional Generative Adversarial Network (GAN) to tackle the problem of image reconstruction in Intensity Interferometry. The proposed GAN-based approach effectively reconstructs the shape, size, and brightness distribution of a fast-rotating star by leveraging sparsely sampled, phase-subtracted Fourier transforms of the source. This method not only enhances the image reconstruction capabilities of Intensity Interferometry but also enables precise parameter estimation in astronomical observations.  
\end{abstract}
\input{intro}
\input{II}
\input{GAN}
\input{ParaGan}
\input{Result}
\input{conclusion}

>>>>

\title{Generative AI for image reconstruction in Intensity Interferometry: a first attempt}
\date{}

@@

\begin{abstract}
In the last few years Intensity Interferometry has made significant strides in achieving high-precision resolution of stellar objects at optical wavelengths. Despite these advancements, phase retrieval remains a major challenge due to the nature of photon correlation. This paper explores the application of a conditional Generative Adversarial Network (cGAN) to tackle the problem of image reconstruction in Intensity Interferometry. This approach successfully reconstructs the shape, size, and brightness distribution of a fast-rotating star from sparsely sampled, spatial power spectrum of the source, corresponding to II with four telescopes.  Although this particular example could also be addressed using parameter fitting, the results suggest that with larger arrays much more complicated systems could be reconstructed by applying machine-learning techniques to II.
\end{abstract}
\input{intro}
\input{II}
\input{GAN}
\input{ParaGan}
\input{Result}
\input{conclusion}
